Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models

### 2020-06-16 analysis/paper/paper.Rmd
dropped a questionable topic sentence (didn't really fit the rest of the paragraph)

<!-- Revise the first sentence of the following paragraph, it seems out of place with the following sentence and paragraph -->

**In fisheries it has long been recognized that both biotic and abiotic factors limit productivity within and across life-stages.** For the purposes of this paper, we define carrying capacity as the maximum number of individuals that can be supported given the quantity and quality of habitat available at a given life-stage. We assume that higher observed relative densities within a given life stage are a function of habitat quantity and quality. Furthermore, we assert that observed fish density is a poor predictor of habitat capacity owing to both a paucity of individuals, especially for threatened or endangered species, and the existence of unmeasured variables that may serve to limit capacity. To address this, we have developed a model to estimate juvenile rearing capacity for Pacific salmon in wadeable streams based on quantile random forest (QRF) [@Meinshausen2006] models using measurements of fish abundance (and density) and habitat characteristics. QRF models combine the theory and justification of quantile regression modeling [@Koenker1978; @Cade2003] with the flexibility and framework of random forest models [@Breiman2001]. They account for unmeasured variables and can be used to describe the entire distribution of predicted fish densities for a given set of habitat conditions, not just the mean expected density. Random forest models have been shown to outperform more standard parametric models in predicting fish-habitat relationships in other contexts [@Knudby2010]. Quantile regression forests share many of the benefits of random forest models, such as the ability to capture non-linear relationships between independent and dependent variables, naturally incorporate interactions between covariates, and work with untransformed data while being robust to outliers [@Prasad2006]. Meanwhile, quantile regression models have been used in a variety of ecological systems to estimate the effect of limiting factors [@Terrell1996; @Cade2003].

>>>>

For the purposes of this paper, we define carrying capacity as the maximum number of individuals that can be supported given the quantity and quality of habitat available at a given life-stage. We assume that higher observed relative densities within a given life stage are a function of habitat quantity and quality. Furthermore, we assert that observed fish density is a poor predictor of habitat capacity owing to both a paucity of individuals, especially for threatened or endangered species, and the existence of unmeasured variables that may serve to limit capacity. To address this, we have developed a model to estimate juvenile rearing capacity for Pacific salmon in wadeable streams based on quantile random forest (QRF) [@Meinshausen2006] models using measurements of fish abundance (and density) and habitat characteristics. QRF models combine the theory and justification of quantile regression modeling [@Koenker1978; @Cade2003] with the flexibility and framework of random forest models [@Breiman2001]. They account for unmeasured variables and can be used to describe the entire distribution of predicted fish densities for a given set of habitat conditions, not just the mean expected density. Random forest models have been shown to outperform more standard parametric models in predicting fish-habitat relationships in other contexts [@Knudby2010]. Quantile regression forests share many of the benefits of random forest models, such as the ability to capture non-linear relationships between independent and dependent variables, naturally incorporate interactions between covariates, and work with untransformed data while being robust to outliers [@Prasad2006]. Meanwhile, quantile regression models have been used in a variety of ecological systems to estimate the effect of limiting factors [@Terrell1996; @Cade2003].
